---
title: "NVIDIA Stock Analysis"
author: "Joshua Agbroko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Overview

This project analyzes NVIDIA's stock performance and compares it to the S&P 500
benchmark using R and Python(reticulate). The objective is to determine whether
an equally weighted, long-only trading strategy based on NVIDIA's relative 
performance can outperform the S&P 500 index. 

```{r message=FALSE, warning=FALSE}
# Loading necessaru libraries
library(tidyverse)
library(httr)
library(jsonlite)
library(ggplot2)
library(lubridate)
library(TTR)
library(reticulate)
```

```{r source Python script }
# Sourcing python script
source_python("fetch_data.py")

# Fetch NVIDIA (NVDA) stock data
nvda_data <- py$get_stock_data("NVDA")

# Converting to R dataframe
nvda_df <- as.tibble(nvda_data)

# View first few rows
print(head(nvda_df))

```

```{r}

```

# Exploratory Data Analysis On NVIDIA Stock Data


```{r Data Validation}
# Check structure 
print(str(nvda_df))

nvda_df$Date <- as.Date(nvda_df$Date)
# Check for missing Values
print(sum(is.na(nvda_df)))


```

```{r}
# Summary statistics
summary(nvda_df)
```
### Stock Price Trends
```{r}
# Line Chart of closing prices
ggplot(nvda_df, aes(x = Date, y = close)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "NVIDIA Stock Price Trend (last 5 Years)",
       x = "Date", y = "Closing Price(USD)") +
  theme_minimal()
```

This shows how NVIDIA's stock has performed over time. We can see a rapid rise in 
closing stock price from 2023 till it eventually stabilizes in early 2025. This should
primarily be due to the immense demand for its high-performance AI chips, which 
are considered to be the industry leader, with its latest Blackwell AI architecture
seeing particularly strong demand and fueling investor confidence in the company's
future potential.


### Daily Returns
This calculates percentage changes in price daily.

```{r Calculaing daily returns}
nvda_df <- nvda_df %>%
  arrange(Date) %>%
  mutate(daily_return = (close / lag(close)) - 1)

print(summary(nvda_df$daily_return))

head(nvda_df)

```


### Volatility Analysis
```{r Histogram of Returns}
ggplot(nvda_df, aes(x = daily_return)) +
  geom_histogram(binwidth = 0.01, fill = "blue",
                 color = "black", alpha = 0.7) +
  labs(title = "Distribution of NVIDIA Daily Returns",
       x = "Daily Return", y = "Frequency") +
  theme_minimal()
```

This shows return distribution, helping us understand volatility. NVIDIA's returns
appear to be normally distributed suggesting that stock is volatile( higher risk, 
but also potential for higher returns)


### Comparison with the S&P 500
 The S&P 500 (Standard & Poor's 500) is a stock market index benchmark that is used
 to track the performance of 500 large U.S. companies across various sectors. It is 
 considered a good representation of the U.S. economy and overall stock market
 performance.

```{r}
# Fetching S&P  500 ETF data 
sp500_data <- py$get_stock_data("SPY")

# converting to R dataframe
sp500_df <- as.tibble(sp500_data)
sp500_df$Date <- as.Date(sp500_df$Date)
  
print(head(sp500_df))
```

```{r Preparing NVIDIA and S&P 500 for comparison}
# Calc. S&P 500 daily returns
sp500_df <- sp500_df %>%
  arrange(Date) %>%
  mutate(daily_return = (close / lag(close)) - 1)

# merging both stock datasets on Date
stock_data <- left_join(nvda_df, sp500_df, by = "Date", 
                        suffix = c("_nvda", "_sp500"))

# Rem. NA values
stock_data <- na.omit(stock_data)

stock_data$Date <- as.Date(stock_data$Date)

# Summary of returns
summary(stock_data$daily_return_nvda)
summary(stock_data$daily_return_sp500)

```

### Comparing Performance with a T-Test

A t-test checks if NVIDIA's daily returns are significantly different from 
S&P 500's.
```{r Perform t-test}
t_test_result <- t.test(stock_data$daily_return_nvda,
                        stock_data$daily_return_sp500,
                        alternative = "greater")

print(t_test_result)

```
Since p < 0.05, we reject the null hypothesis, meaning NVIDIA's mean returns are
significantly greater than the S&P 500's. The confidence interval suggests that
even in the worst-case scenario, NVIDIA still uotperforms the S&P 500 on average.


```{r Visualizing Performance}
# line plot of NVIDIA vs S&P 500 Returns
ggplot(stock_data, aes(x = Date)) +
  geom_line(aes(y = daily_return_nvda, color = "NVIDIA"), size = 1) +
  geom_line(aes(y = daily_return_sp500, color = "S&P 500"), size = 1) +
  labs(title = "NVIDIA vs S&P 500 Daily Returns",
       x = "Date", y = "Daily Return") +
  scale_color_manual(values = c("NVIDIA" = "blue", "S&P 500" = "red")) +
  theme_gray()

```

This visually compares NVIDIA's performance with the market

## Implementating a Simple Trading Strategy

We will
- Define a trading rule: invest i NVIDIA only if its return is greater than
S&P 500 on a given day
- Track portfolio performance over time
- Compare results to simply holding the S&P 500

### Define the Trading Rule
 We create a binary investment signal:
 - 1(Invest in NVIDIA) -> If daily_return_nvda > daily_return_sp500.
 - 0(Stay in cash) -> Otherwise.
```{r}
# adding  atrading signal column (1 - invest, 0 = no investment)
stock_data <- stock_data %>%
  mutate(trade_signal = ifelse(daily_return_nvda > daily_return_sp500, 1, 0))
# Check
head(stock_data)


```
This identifies the dqays when we invest in NVIDIA.

### Calculating portfolio returns

We assume an equally weighted long-only portfolio and calculate cumulative returns.
```{r}
# Compute portfolio returns
stock_data <- stock_data %>%
  mutate(portfolio_return = trade_signal * daily_return_nvda)

# Compute cumulative returns
stock_data <- stock_data %>%
  mutate(cumulative_portfolio = cumprod(1 + portfolio_return),
         cumulative_sp500 = cumprod(1 + daily_return_sp500))

# Check
tail(stock_data)

```
This tracks how our trading strategy performs compared to the S&P 500.

### Visualizing performance
```{r compare NVIDIA stategy to S&P 500 over time}

ggplot(stock_data, aes(x = Date)) +
  geom_line(aes(y = cumulative_portfolio, color = "NVIDIA Strategy"), size = 1) +
  geom_line(aes(y = cumulative_sp500, color = "S&P 500 Benchmark"), size = 1) +
  labs(title = "Cumulative Return: NVIDIA Strategy vs S&P 500",
       x = "Date", y = "Cumulative Return") +
  scale_color_manual(values = c("NVIDIA Strategy" = "blue", "S&P 500 Benchmark" = "red")) +
  theme_gray()
  
```


### Evaluating Strategy Performance
We compare:
- Average daily return of our strategy vs S&P 500
- Standard deviation (risk)
- Sharpe Ratio(risk-adjusted return)
 
```{r}
# Computing summary statistics 
strategy_mean_return <- mean(stock_data$portfolio_return, na.rm = TRUE)
sp500_mean_return <- mean(stock_data$daily_return_sp500, na.rm = TRUE)

strategy_sd <- sd(stock_data$portfolio_return, na.rm = TRUE)
sp500_sd <- sd(stock_data$daily_return_sp500, na.rm = TRUE)

strategy_sharpe <- strategy_mean_return / strategy_sd
sp500_sharpe <- sp500_mean_return / sp500_sd

# Print
cat("Strategy Mean Return:", strategy_mean_return, "\n")
cat("S&P 500 Mean Return:", sp500_mean_return, "\n")
cat("Strategy Sharpe Ratio:", strategy_sharpe, "\n")
cat("S&P 500 Sharpe Ratio:", sp500_sharpe, "\n")

```

# Conclusion
This project set out to analyze NVIDIA's sock performance relative to S&P  500
index and evaluate whether an equally weighted, long-only investment strategy in
NVIDIA could outperform the benchmark.

Through exploratory data analysis(EDA), return calculations, and statistical
testing(t-test)we found that:

- NVIDIA consistently outperformed the S&P 500 in terms of average daily returns
(1.34% vs 0.0659%).
- A trading strategy that invests in NVIDIA when it beats the S&P 500 achieved 
higher cumulative returns.
- Risk-adjusted performance was superior, as reflected in the higher Sharpe
ratio( 0.592 vs 0.508).
- However, he strategy exhibited higher volatility,meaning that while returns
were higher, the fluctuations in price were more significant than those  of the 
S&P 500. 
 
## Key Takeaways
- NVIDIA outperformed the S&P 500, making it an attractive stock for active 
investment.
- Higher returns come with higher risk, requiring risk management strategies such as
stop-losses or portfolio diversification.
- Future research could expand this approach to assess whether NVIDIA's 
outperformance is consistent or part of a larger market trend.
















